---
title: "5.2.4 Exercises"
author: "Berfin Aş"
date: "09 02 2022"
output: html_document
---

```{r setup, include=FALSE}
library(nycflights13)
library(tidyverse)
library(dbplyr)
knitr::opts_chunk$set(echo = TRUE)
```

## Chapter 5 Section 5.2.4

## Question 1

Find the flights that

1. Had an arrival delay of two or more hours

```{r}
filter(flights, arr_delay >= 120)
```

2. Flew to Houston (IAH or HOU)

```{r}
filter(flights, dest == "IAH" | dest == "HOU")
(flew_houston <-  filter(flights, dest %in% c("IAH", "HOU")))
```

3. Were operated by United, American, or Delta

```{r}
filter(flights, carrier %in% c("UA", "DL", "AA"))
```

4. Departed in summer (July, August, and September)

```{r}
filter(flights, month %in% c(7,8,9))
```

5. Arrived more than two hours late, but didn’t leave late

```{r}
filter(flights, arr_delay > 120, dep_delay <= 0)
```

6. Were delayed by at least an hour, but made up over 30 minutes in flight

```{r}
filter(flights, dep_delay >= 60, flight > 30)
```

7. Departed between midnight and 6am (inclusive)


```{r}
hms <- format(as.POSIXct(flights$time_hour), "%H:%M:%S" )

dep_midto6 <- flights[hms >= "00:00:00" & hms <= "06:00:00",]
dep_midto6

```

or we can write it by considering the fact that midnight in dep_time is 2400

```{r}
summary(flights$dep_time)
filter(flights, dep_time <= 600 | dep_time ==2400 ) 
```

## Question 2

Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?

Answer: between(x, left, right) is equivalent to x >= left & x <= right. We can rewrite the 4th part of the first question (Departed in summer (July, August, and September)) as;

```{r}
filter(flights, between(month,7,9))
```

## Question 3

How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

```{r}
filter(flights, is.na(dep_time))
```

It can be seen that arr_time, dep_delay, arr_delay and air_time are also missing for these rows.

## Chapter 5 Section 5.3.1

## Question 2

Sort flights to find the most delayed flights. Find the flights that left earliest.

```{r}
arrange(flights, desc(dep_delay))
arrange(flights, dep_delay)
```

## Question 3

Sort flights to find the fastest (highest speed) flights.

```{r}
arrange(flights, desc(distance/air_time))
```


## Question 4

Which flights travelled the farthest? Which travelled the shortest?

```{r}
arrange(flights, desc(distance))
arrange(flights, - desc(distance))
```

## Chapter 5 Section 5.6.7

## Question 5

Which carrier has the worst delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about flights %>% group_by(carrier, dest) %>% summarise(n()))

We can compare the airports and carriers by finding their max delays and then comparing it to means. We can then make conclusions by looking at ff table

```{r}
flights %>%
  group_by(carrier)%>%
  summarise(
    max_arr = max(arr_delay, na.rm= TRUE),
    max_dep = max(dep_delay, na.rm =TRUE)) %>%
  arrange(desc(max_arr, max_dep)) %>%
  filter(1:n() == 1)
                   

```

```{r}
flights%>%
  group_by(carrier)%>%
  summarise(
    mean_arr = mean(arr_delay,na.rm= TRUE),
    mean_dep = mean(dep_delay,na.rm= TRUE))%>%
  arrange(desc(mean_arr,mean_dep))%>%
  filter(1:n() == 1)


```

```{r}
ff <- flights%>%
  group_by(carrier,origin,dest)%>%
  summarise(
    mean_arr = mean(arr_delay,na.rm= TRUE),
    mean_dep = mean(dep_delay,na.rm= TRUE))%>%
  arrange(desc(mean_arr,mean_dep))%>%
  filter(1:n() == 1)

```

```{r}
view(ff)
```


## Chapter 5 Section 5.7.1

## Question 2

Which plane (tailnum) has the worst on-time record?

```{r}
flights_sml <-  select(flights,
  year:day,
  ends_with("delay"),
  distance,
  air_time
)
```


```{r}
flights_sml %>%
  group_by(year, month,day) %>%
  filter(rank(desc(arr_delay)) < 10)

```


```{r}
flights %>%
filter(!is.na(tailnum)) %>%
  mutate(on_time = !is.na(arr_time) & (arr_delay <= 0)) %>%
  group_by(tailnum) %>%
  summarise(on_time = mean(on_time), n = n()) %>%
  filter(min_rank(on_time) == 1)

```


## Question 3

What time of day should you fly if you want to avoid delays as much as possible?

```{r}
flights%>%
  group_by(hour)%>%
  summarise(arr_delay= mean(arr_delay,na.rm= TRUE))%>%
  arrange(arr_delay)
```

